---
title: "Exploratory Data Analysis"
author: "Aymen Belakhoua"
date: "5/10/2022"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

```
# Final Report

## GWEI Prices over Time

------

## Import Necessary Packages
```{r echo=T, results='hide'}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(ggpubr)
```

## Read CSV

```{r pressure, echo=FALSE}
setwd(r"(C:\Users\Aymen\Desktop\Intro DS - 3010\Final\Completed)")
df <- read.csv('GasGot.csv')
head(df, 5)
```

## Change Datetime to Hour

```{r}
df$datetime <- format(as.POSIXct(df$datetime,
                                 format="%Y-%m-%d %H:%M:%S",
                                 tz="UTC"),
                      format = "%H")
names(df)[names(df)== 'datetime'] <- "hour"
df <- rename_with(df, toupper)
head(df, 5)
```

------

## Plot Density Plot

```{r}
ggplot(df, aes(x=AVERAGE)) + 
  geom_histogram(aes(y=..density..), colour="blue", fill="white", binwidth = 10)+
  geom_density(alpha=.2, fill="#FF6666") +
  xlim(c(0,200)) + 
  xlab("Average") +
  ylab("Density")
```

The histogram is right skewed, meaning the user will pay more than the mode (most common) price on average.

-----

### Get Mean Median Mode and Standard Deviation of "Average" Gas Price.

```{r}
getmode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}
```

The calculations for the 'Average' Speed's cost for the entire data set are as follows-

* Mean: `r round(mean(df$AVERAGE))`

* Median: `r median(df$AVERAGE)`

* Mode: `r getmode(round(df$AVERAGE))`

* STD: `r round(sd(df$AVERAGE))`

```{r}
df2 <- read.csv('GasGot.csv')
df2$datetime <- as.POSIXct(df2$datetime,
                                 format="%Y-%m-%d %H:%M:%S",
                                 tz="UTC")

#df2$datetime <- as.numeric(df2$datetime)
cor(x=as.numeric(df2$datetime), y=df2$average,  method = "pearson", use = "complete.obs")

ggscatter(df2, x = 'datetime', y = 'average', 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Date", ylab = "Average Cost\nGWEI")
```

The trend line is sloped upwards, however there is a very low linear correlation (.161) so this is not indicative of much.")

-----

## Hourly STD Calculations and Visualization

```{r}
hours = c()
std = c()
for (i in 0:23){
  hours <- c(hours, i)
  df2 <- filter(df, as.integer(HOUR) == i)
  deviation = sd(df2$AVERAGE)
  std <- c(std, deviation)
}
dfSTD <- data.frame(hours, std)
head(dfSTD)

ggplot(dfSTD, aes(x = hours, y= std)) + geom_point() +
  geom_hline(yintercept = mean(std))
```

Average STD: `r round(mean(dfSTD$std))`

There is a significantly high deviation in GWEI Price.

-----

### Average GWEI Price Boxplot plotted hourly, ordered by median value, with outliers removed.

```{r}
ggplot(df, aes(x=reorder(HOUR, AVERAGE), y=AVERAGE)) +
  geom_boxplot(outlier.shape = NA) +
  labs(y="Average Speed\n(GWEI)", x="Hour", 
       title="Boxplot: Hours Sorted by Mean") +
  ylim(0,250)
```

-----

## Hourly Linear Correlation

```{r}
df$HOUR <- factor(df$HOUR)

averages <- c()
hours <-c()
for (hour in unique(df$HOUR)){
  hours <- append(hours, hour)
  averages <- append(x=averages,
                     value = mean(df[df$HOUR == hour,'AVERAGE']))
}
df3 <- list(average = averages, hours = hours)
df3 <- as.data.frame(df3)
df3 <- df3[order(averages),]

ggscatter(df3, x = "hours", y = 'average', 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Hours", ylab = "Average")
```

### Correlation Test

```{r}
cor.test(as.numeric(df3$hours), df3$average, method=c("pearson", "kendall", "spearman"))
```


With a correlation coefficient of >.5 the model indicates that there is a significant linear association.

-----

## Hourly Linear Regression

```{r}
model <- lm(as.numeric(average)~as.numeric(hours), df3)
summary(model)
```

With an adjusted R-Squared of **.2734** we can see that the hour of the day has a weak effect (low correlation) on the GWEI cost.

With a p value of **.0051** we can confidently reject the null hypothesis.